---
title: "Homework7"
author: "Tracie-Lynn Lamoreux, Macy Schanbacher, Nic Rhoads"
date: "2023-03-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggplot2)
library(ggmosaic)
HAPPY<- readRDS("Happy.rds")
```

Cleaning:
```{r}
HAPPY[HAPPY == "DK" | HAPPY == "IAP" | HAPPY == "NA"] <- NA
HAPPY$AGE[HAPPY$AGE == "89 OR OLDER"] <-89
str(HAPPY)
```

Casting columns:
```{r}
HAPPY$HAPPY <- as.factor(HAPPY$HAPPY)
HAPPY$YEAR <- as.numeric(HAPPY$YEAR)
HAPPY$AGE <- as.numeric(HAPPY$AGE)
HAPPY$SEX <- as.factor(HAPPY$SEX)
HAPPY$MARITAL <- as.factor(HAPPY$MARITAL)
HAPPY$DEGREE <- as.factor(HAPPY$DEGREE)
HAPPY$FINRELA <- as.factor(HAPPY$FINRELA)
HAPPY$HEALTH <- as.factor(HAPPY$HEALTH)
HAPPY$PARTYID <- as.factor(HAPPY$PARTYID)
HAPPY$POLVIEWS <- as.factor(HAPPY$POLVIEWS)
```

Re-ordering levels:
```{r}
HAPPY$DEGREE <- factor(HAPPY$DEGREE, levels = levels(reorder(HAPPY$DEGREE, HAPPY$AGE, na.rm = TRUE)))
HAPPY$MARITAL <- factor(HAPPY$MARITAL, levels = levels(reorder(HAPPY$MARITAL, HAPPY$AGE, na.rm = TRUE)))
HAPPY$HAPPY <- factor(HAPPY$HAPPY, levels = c("NOT TOO HAPPY", "PRETTY HAPPY", "VERY HAPPY"))
HAPPY$FINRELA <- factor(HAPPY$FINRELA, levels = c("FAR BELOW AVERAGE", "BELOW AVERAGE", "AVERAGE","ABOVE AVERAGE", "FAR ABOVE AVERAGE"))
HAPPY$HEALTH <- factor(HAPPY$HEALTH, levels = c("POOR", "FAIR", "GOOD", "EXCELLENT"))
```

Tracie-Lynn's question:
```{r}
ggplot(data = HAPPY) + geom_mosaic(aes(x = product(MARITAL), fill = HAPPY, weight = 1), na.rm = TRUE)+ facet_wrap(HAPPY$HEALTH) + theme_classic()
```
Tracie-Lynn decided to investigate the relationship of happiness, marital status, and health. It seems like a 
lot more people are happy if they are in better health and married as opposed to poor health and never married married.  
Macy's question: Macy's question: Happiness, Degree, Finrela

```{r}
library(ggplot2)
library(ggmosaic)
ggplot(HAPPY) + geom_mosaic(aes(x = product(DEGREE), fill = HAPPY, weight = 1), na.rm = TRUE) + facet_wrap(~FINRELA)
```

Nic's question: I looked at the relationship between happiness, marital status, and sex. People who are married are notably more likely to be very happy and are less likely to be not too happy. People who are separated have the lowest happiness rate and the highest happiness rate. Both of these carry between the sexes. Males, regardless of marital status, are less likely to be very happy, but they have similar rates of not too happy.

```{r}
ggplot(data = HAPPY) + geom_mosaic(aes(x = product(MARITAL), fill = HAPPY, weight = 1), na.rm = TRUE) + facet_wrap(HAPPY$SEX)
```
